I.Sicily. An EpiDoc Corpus for Ancient Sicily

### 2016-02-09 source/31/EAGLE2016_I Sicily_final copy.tex
corrections to Prag

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{I.Sicily. An EpiDoc corpus for ancient Sicily}
\author[OXF]{Jonathan Prag\corref{first}}
\author[opensky]{James Chartrand}
\author[OXF]{James Cummings}
\address[OXF]{University of Oxford}
\address[opensky]{OpenSky Solutions}
\cortext[first]{Corresponding author. Email: jonathan.prag@merton.ox.ac.uk}
\date{2015-11-18}
\begin{document}

@@

\section{Introduction: What is \emph{I.Sicily}}
\noindent \emph{I.Sicily} is an online, open access, digital corpus of the inscriptions on stone from ancient Sicily.\footnote{ The
corpus will be mounted at \url{www.csad.ox.ac.uk/sicily/isicily}, but is presently on a development server. A public face is
currently maintained via a blog at \url{http://isicily.wordpress.com/}, as well as on Facebook at \url{www.facebook.com/ISicily}
and on Twitter via \url{I.Sicily@Sicilyepigraphy}.} The corpus aims to include all texts inscribed on stone, in any language,
between approximately the seventh century BC and the seventh century AD. The corpus currently contains records for over
2,500 texts, and when complete is likely to contain c.4,000. The corpus is built upon a conversion from a legacy
dataset of metadata in MS Access to EpiDoc TEI XML.\footnote{\url{http://sourceforge.net/p/epidoc/wiki/Home/} [accessed
26.09.2015].} The XML records are held in an eXist database for xQuery access, and additionally indexed for full-text
search using SOLR/Lucene. The corpus and related information (museum list, bibliography) are published as Linked Data,
and are manipulated through a RESTful API. The records are queried and viewed through a web interface built with
AngularJS and jQuery javascript components. Mapping is provided in the browser by the Google Maps API, and ZPR (Zoom,
Pan, Rotate) image-viewing is provided by the IIP image server and the OpenSeadragon javascript library.

At the time of writing (September 2015), the main conversion routine is being refined and the epigraphic texts are being
collated for incorporation into the records. An ancillary database of museum collections in Sicily has been constructed
and bibliography is held in a Zotero library. Extensive search facilities will be provided, including map-based and
bibliographic searching. Individual inscriptions and individual museums will both be provided with URIs, as will
personal names and individuals; places will be referenced using \emph{Pleiades}; epigraphic types, materials, and supports
using the EAGLE vocabularies.

\section{The motivations for and origins of \emph{I.Sicily}}
\noindent The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth century tradition, the existing \emph{copora} are far from complete and the ability of key journals such as \emph{SEG}
or \emph{AE} to keep pace with local publication has been limited.\footnote{ For an overview of the \emph{corpora} tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based \emph{corpora} have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

@@

\noindent Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the \emph{Corpus Inscriptionum Iudaeae/Palaestinae} (\emph{CIIP}), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{Cotton:2007aa}.} \emph{I.Sicily} sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific \emph{corpora} is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The \emph{Lexicon of Greek
Personal Names} records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no \emph{onomasticon} for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, \emph{I.Sicily} will generate a
powerful tool for future study. Although \emph{I.Sicily} in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these text in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See \citet{prag_epigraphy_2002} for a first effort in this direction.}

@@

\url{http://csad.ox.ac.uk/sicily/inscription/ISic1234}

@@

The richness of \emph{I.Sicily}’s records in this area mean that \emph{I.Sicily} is currently collaborating with both \emph{Trismegistos} and
\emph{IDES} (`Integrating Digital Epigraphies').\footnote{\url{www.trismegistos.org/} [accessed 26.09.2015] and \url{http://ides.io/}
[accessed 26.09.2105].} The former aims to generate \emph{TM} numbers for all the Sicilian material (which \emph{I.Sicily} will
include); the latter is to assist \emph{IDES} in the refining of links between, e.g., \emph{PHI} and \emph{SEG} records, and to improve
\emph{I.Sicily}’s own recording of PHI numbers.

@@

\url{http://csad.ox.ac.uk/sicily/museum/SicMus123}

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{I.Sicily. An EpiDoc Corpus for Ancient Sicily}
\author[OXF]{Jonathan Prag\corref{first}}
\author[opensky]{James Chartrand}
\author[OXF]{James Cummings}
\address[OXF]{University of Oxford}
\address[opensky]{OpenSky Solutions}
\cortext[first]{Corresponding author. Email: jonathan.prag@merton.ox.ac.uk}
\date{2015-11-18}
\begin{document}

@@

\section{Introduction: What is \emph{I.Sicily}}
\noindent \emph{I.Sicily} is an online, open access, digital corpus of the inscriptions on stone from ancient Sicily.\footnote{ The
corpus will be mounted at \url{www.sicily.classics.ox.ac.uk/isicily/}, but is presently on a development server. A public face is
currently maintained via a blog at \url{http://isicily.wordpress.com/}, as well as on Facebook at \url{www.facebook.com/ISicily}
and on Twitter via \url{I.Sicily@Sicilyepigraphy}.} The corpus aims to include all texts inscribed on stone, in any language,
between approximately the seventh century BC and the seventh century AD. The corpus currently contains records for over
2,500 texts, and when complete is likely to contain c.4,000. The corpus is built upon a conversion from a legacy
dataset of metadata in MS Access to EpiDoc TEI XML.\footnote{\url{http://sourceforge.net/p/epidoc/wiki/Home/} [accessed
26.09.2015].} The XML records are held in an eXist database for xQuery access, and additionally indexed for full-text
search using SOLR/Lucene. The corpus and related information (museum list, bibliography) are published as Linked Data,
and are manipulated through a RESTful API. The records are queried and viewed through a web interface built with
AngularJS and jQuery javascript components. Mapping is provided in the browser by the Google Maps API, and ZPR (Zoom,
Pan, Rotate) image-viewing is provided by the IIP image server and the OpenSeadragon javascript library.

At the time of writing (September 2015), the main conversion routine is being refined and the epigraphic texts are being
collated for incorporation into the records. An ancillary database of museum collections and archaeological sites in Sicily has been constructed
and bibliography is held in a Zotero library. Extensive search facilities will be provided, including map-based and
bibliographic searching. Individual inscriptions and individual museums will both be provided with URIs, as will
personal names and individuals; places will be referenced using \emph{Pleiades}; epigraphic types, materials, and supports
using the EAGLE vocabularies.

\section{The motivations for and origins of \emph{I.Sicily}}
\noindent The existing epigraphic landscape in Sicily is extremely diverse in two primary regards: on the one hand, the island has
a very mixed cultural and linguistic make-up, meaning that the epigraphic material is itself extremely varied, with
extensive use throughout antiquity of both Greek and Latin, as well as Oscan, Punic, Sikel, and Hebrew;\footnote{
Recent overview of much of the linguistic tradition in \citet{tribulato_language_2012}; and of the epigraphic material in \citet{gulletta_sicilia_1999}.} on the other hand, the publication of this material has a very uneven record and despite an excellent
pre-twentieth-century tradition, the existing \emph{copora} are far from complete and the ability of key journals such as \emph{SEG}
or \emph{AE} to keep pace with local publication has been limited.\footnote{ For an overview of the \emph{corpora} tradition up to
the twentieth century, see \citet{de_vido_corpora_1999}.} A limited number of museum-based \emph{corpora} have been published in recent
decades (for Catania, Palermo, Messina, and Termini Imerese, as well as the material from Lipari), but this has not
greatly improved the overall situation.

@@

\noindent Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the \emph{Corpus Inscriptionum Iudaeae/Palaestinae} (\emph{CIIP}), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{Cotton:2007aa}.} \emph{I.Sicily} sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific \emph{corpora} is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The \emph{Lexicon of Greek
Personal Names} records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no \emph{onomasticon} for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, \emph{I.Sicily} will generate a
powerful tool for future study. Although \emph{I.Sicily} in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these texts in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See \citet{prag_epigraphy_2002} for a first effort in this direction.}

@@

\url{sicily.classics.ox.ac.uk/isicily/inscriptions/1234}

@@

The richness of \emph{I.Sicily}’s records in this area means that \emph{I.Sicily} is currently collaborating with both \emph{Trismegistos} and
\emph{IDES} (`Integrating Digital Epigraphies').\footnote{\url{www.trismegistos.org/} [accessed 26.09.2015] and \url{http://ides.io/}
[accessed 26.09.2105].} The former aims to generate \emph{TM} numbers for all the Sicilian material (which \emph{I.Sicily} will
include); the latter is to assist \emph{IDES} in the refining of links between, e.g., \emph{PHI} and \emph{SEG} records, and to improve
\emph{I.Sicily}’s own recording of \emph{PHI} numbers.

@@

\url{sicily.classics.ox.ac.uk/isicily/museums/123}
